---
title: "30-feature-engineering"
output: html_notebook
---
The first thing to do here would be calculate the scores for the big five dimensions.

Calculation the Extraversion, Aggreableness, Neuroticism, Conscientiousness and Openness to experience, scores. 
The score determine which range you are in.

The test has 10 questions for each trait. (For Extraversion - E1:E10, Agreeableness - A1:A10, etc) Each question has a five point scale going from "0 -  strongly disagree" to "5 -  strongly agree".

These questions would either favor the specfic personality trait, or go against it. In my analysis, I chose to combine the scores from each dimension into a single dimension and then added or substracted the score from a single question, based on if the question supports or goes against a particular trait. 

for example:

E1	I am the life of the party. - Score from this question would be added
E2	I don't talk a lot. - Score from this question would be substracted

#Creating dimensional scores
```{r Creation of Dimensional Score}

get_scores <- function(df){
  #Add asserts here to check if all are not na. If there are NA's, remove them. 
  #[ADD CODE HERE]
  df %>% 
    mutate(extraversion = E1 - E2 + E3 - E4 + E5 - E6 + E7 - E8 + E9 - E10, 
           neuroticism = N10 + N9 + N8 + N7 + N6 + N5 - N4 + N3 - N2 + N1, 
           agreeableness = A10 + A9 + A8 - A7 + A6 - A5 + A4 - A3 + A2 - A1, 
           conscientiousness = C10 + C9 - C8 + C7 - C6 + C5 - C4 + C3 - C2 + C1, 
           openness = O10 + O9 - O8 + O7 - O6 + O5 - O4 + O3 - O2 + O1) %>% 
    select(extraversion, neuroticism, agreeableness, conscientiousness, openness,  everything())
}

#Verify that there are no NA's
personas %>% 
  get_scores() %>% 
  verify(!is.na(c("extraversion", "neuroticism", "agreeableness", "conscientiousness", "openness")))

```

#Normalization
We will need to normalize these as all of them have different ranges. 

```{r Normalization By Variable Function}
normalize <- function(x)(x -min(x, na.rm = T))/(max(x, na.rm = T)-min(x, na.rm = T))
norm_by_var <- function(df, vars){
  df %>% 
    mutate_at(vars, list(norm = normalize)) %>% 
    select(vars, contains('norm'), everything())
}

#Verifies normalized values created are within the bounds of 0 and 1

personas%>% 
  dimensional_scores() %>% 
  norm_by_var(c("extraversion","neuroticism", "agreeableness","conscientiousness", "openness")) %>% 
  assert(within_bounds(0,1), contains("norm"))

```

#Creating age brackets

We will need to split up the age in a way that is reasonable. Let's first look at the distribution of the age of the people who took the personality assessment. 

```{r Creating an age bracket}
personas%>% 
  ggplot(aes(x = age))+
  geom_histogram()

personas%>% 
  group_by(age) %>% 
  summarize(n=n()) %>% 
  ungroup() %>% 
  arrange(age, -n)
```
From here we can see that the number of observations or subjects of age 53 and older is going to be lower than 100. I think this will reduce the quality of summary measures that we can derive from the splits.
For this reason, I will cut off the ages past 60. 

```{r}
personas%>% 
  dimensional_scores() %>% 
  norm_by_var(c("extraversion","neuroticism", "agreeableness","conscientiousness", "openness")) %>% 
  filter(age<=60)

```


The brackets I will create will be:
Bracket 1: 13-18 - Teenasge years
Bracket 2: 19-21 - Early adulthood
Bracket 3: 22-25 - Early 20s
Bracket 4: 26-29 - later 20s
Bracket 5: 30-34 - Early 30s
Bracket 6: 35-39 - Late 30s
Bracket 7: 40-44 - Early 40s
Bracket 8: 45-49 - Late 40s
Bracket 9: 50-60 - 50s 



```{r Age Bracket Function}

bracket_creator <- function(x){
  if (x >= 13 & x <= 18){"Bracket1"}
  else if (x >= 19 & x <= 21){"Bracket2"}
  else if (x >= 22 & x <= 25){"Bracket3"}
  else if (x >= 26 & x <= 29){"Bracket4"}
  else if (x >= 30 & x <= 34){"Bracket5"}
  else if (x >= 35 & x <= 39){"Bracket6"}
  else if (x >= 40 & x <= 44){"Bracket7"}
  else if (x >= 45 & x <= 49){"Bracket8"}
  else {"Bracket9"}
}

bracket_creator <- Vectorize(bracket_creator)

add_age_bracks <- function(df){
  df %>% 
    mutate_at("age", list(brackets = bracket_creator)) %>% 
    select("age", brackets, everything())
}

#Verify ages are within set
personas%>% 
  dimensional_scores() %>% 
  norm_by_var(c("extraversion","neuroticism", "agreeableness","conscientiousness", "openness")) %>% 
  add_age_bracks() %>% 
  assert(in_set("Bracket1","Bracket2","Bracket3", "Bracket4", "Bracket5", "Bracket6", "Bracket7","Bracket8", "Bracket9"), brackets)
  

```

#Standardizing scores based on age brackets
This will give us a better estimation of where each person stands in those dimension with respect to scores of those around their age. 
```{r}
scale2 <- function(x, na.rm = T)(x -mean(x, na.rm = na.rm))/sd(x, na.rm = na.rm)

mean_by_bracket <- function(df, vars){
  df %>% 
    group_by(brackets) %>% 
    mutate_at(vars, list(std_brack = scale2)) %>% 
    ungroup() %>% 
    select(age, brackets, contains('std_brack'), everything())
}

#Verify it is not NA
personas%>% 
  dimensional_scores() %>% 
  norm_by_var(c("extraversion","neuroticism", "agreeableness","conscientiousness", "openness")) %>% 
  add_age_bracks() %>% 
  mean_by_bracket(c("extraversion","neuroticism", "agreeableness","conscientiousness", "openness")) %>% 
  verify(!is.na(contains("std_brack")))
```

