---
title: "assign4"
output: html_notebook
---


For the regular expression exercises, try writing first using regular expressions directly, then try using rverbalexpressions. Provide the code for both. 

```{r}
library(lubridate)
library(tidyverse)
library(stringr)
library(RVerbalExpressions)
```

# Part 1

## R4DS 16.4.5

**Create a vector of dates giving the first day of every month in the current year.**

```{r}
day_1_months <- seq(ymd('2019-01-01'),ymd('2019-12-01'),by='months')
day_1_months

```


**Write a function that given your birthday (as a date), returns how old you are in years.**

```{r}
age <- function(bday){
  bday <-ymd(bday)
  return(interval(bday,today())/years(1))
}

age("1994-11-11")

```


**Write a function that given your birthday, returns the day of the week you were born on.**

```{r}
#Sat,Sun,Mon,Tue,Wed,Thu,Fri

birth_day <- function(bday){
  week_days = c("Saturday","Sunday","Monday", "Tuesday","Wednesday","Thursday","Friday")
  return(week_days[wday(ymd(bday))])
}


birth_day("1994-11-11")

```


## R4DS 14.3.2.1

Given the corpus of common words in stringr::words, create regular expressions that **find all words that have seven letters or more.** (Since this list is long, you might want to use the match argument to str_view() to show only the matching or non-matching words.)

```{r}

words<- stringr::words


#string
str_view(words, ".......", match = T)

```



## R4DS 14.3.3.1

Use str_detect to find all words that:

**That only contain consonants. (Hint: thinking about matching “not”-vowels.)**

```{r}

words[str_detect(string = words, pattern = "^[^aeiou]+$")]

```



**Empirically verify the rule “i before e except after c”.**
```{r}

words[str_detect(string = words, pattern = "cei|[^c]ie")]

```


**Create a regular expression that will match telephone numbers as commonly written in the US. Split this up into multiple expressions as needed.**

```{r}
#input vector contains numbers entered as 

input <- c("217-377-2867", "217.377.2867", "EM7.377.2867", "XYX-234-5467", "666.768-7890")
bool <-str_detect(input, "^\\d{3}[-\\.]\\d{3}[-\\.]\\d{4}$")
input[bool]

```

# Part 2


Choose a dataset (this can be the same dataset as the last assignment, or a new one). Define a predictive modeling problem. Create the appropriate files and notebooks to do the following:

1. Describe the data and the problem

2. Read in and check data

2. Clean up the data. 

*Note: You may request that score for these sections replace your score for the coding portion of Assignment 3.*

4. Using best practices, write four functions which add engineered features to the dataset, including a description and rationale. Include a test to determine whether the features are working as intended. 

5. Prepare the data for modeling

Note: this will form the basis for your midterm project. Take advantage of TA's and my office hours. We can provide feedback and guidance. 

